---
output:
  pdf_document: default
  html_document: default
---

Group ID: Group 1

Member1 : Vilok C Bhat 
USN     : 1NT20IS187

Member2 : Byalalli Jagadish
USN     : 1NT20IS037


```{r}
library(dplyr)        # For data manipulation
library(ggplot2)      # For data visualization
library(caret)        # For machine learning functions

```

```{r}
# Assuming the dataset is in CSV format
dataset <- read.csv("D:/Download/gender_submission.csv", stringsAsFactors = FALSE)

```

```{r}
getwd()
str(dataset)
```

```{r}
head(dataset)
```

```{r}
tail(dataset)
```

```{r}
summary(dataset)
```
```{r}
sum(dataset)
```
```{r}
train <- read.csv("D:/Download/train.csv", stringsAsFactors = F, header = TRUE)
```

```{r}
head(train)
str(train)
```
```{r}
ls()
ls(pattern = "d")
```
pattern match
```{r}
ls(pattern='^t')
```
```{r}
ls(pattern='a$')
```
```{r}
ls(pattern = '^b|^e')
```
```{r}
ls(pattern='e.a')
```
deleting the list 
```{r}
rm(fw)
```

```{r}
test <- read.csv("D:/Download/test.csv", stringsAsFactors = F)
head(test)
str(test)
```
```{r}
test
```
```{r}
colnames(train)
```
```{r}
row.names(train)
```
```{r}
class(test)
```
```{r}
temp1=as.character(train$Survived)
class(temp1)
```
```{r}
temp2=as.integer(temp1)
class(temp2)
temp2

```
```{r}
temp3=as.numeric(temp2)
class(temp3)
```
```{r}
temp4=as.factor(temp3)
class(temp4)
```
```{r}
dataset[1]
```
```{r}
ls.str(pattern = 'data')
```
history save
#```{r}
#savehistory(file = '.Rhistory')
#loadhistory(file='.Rhistory')
#```

```{r}
max(dataset)
```
```{r}
which(dataset==max(dataset))
```
generating sequence
```{r}
dataset[seq(1, length(dataset), 2)]
```
```{r}
temp=c("sec a","sec b","sec a","sec b","sec a","sec c","sec c")
sort(temp)
```
```{r}
sort(temp,decreasing = TRUE)
```

```{r}
rank(temp)
```
```{r}
order(temp)
```
cumulative sum
```{r}
cumsum(temp3)
```
complex cumulative record
```{r}
seq(temp)
```
The apply() Command for Summaries on Rows or Columns
```{r}
margin.table(train$Age,1)
```
```{r}
mean(train$PassengerId)
```
```{r}
median(train$Pclass)
```
```{r}
is.matrix(train)
```
Stem and Leaf Plot
```{r}
stem(train$Age)
```
```{r}
hist(train$Pclass, col='gray75', main=NULL, xlab = 'pclass',
ylim=c(0, 0.3), freq = FALSE)
```
```{r}
names(train)
```
```{r}
plot(train$Pclass)
```

```{r}
runif(10)
```
runif()creates ten random numbers;
```{r}
runif(10)
```
punif is to create probability of 10 numbers
```{r}
punif(6, min = 0, max = 10)
```




```{r}
# bind training & test data
full  <- bind_rows(train, test) 
str(full)
dim(full)
```

The ratio of survived
```{r}
# check the number of survived(=1) in train data
table(train$Survived)
addmargins(table(train$Survived))

# and ratio of survived(=1) in train data
prop.table(table(train$Survived))
addmargins(prop.table(table(train$Survived)))
```
```{r}
# add Survived column into test with 418 non-survidved (=0)
str(test)
test$Survived <- rep(0, 418)
dim(test)
# that's a totally blind guess by making everyone pass away, but it is a simple tactics
```

```{r}
# make the first submission with R, rate = 0.622
submit <- data.frame(PassengerId = test$PassengerId, Survived = test$Survived)
write.csv(submit, file = "theyallgone.csv", row.names = FALSE)
```

The ratio of survived by sex and age
Getting the sex info
```{r}
# check the number of passegers by sex in train data
table(train$Sex) # plain numbers
addmargins(table(train$Sex)) # plain numbers with control sum (1.000000)

# and ratio of passengers by sex in train data
prop.table(table(train$Sex)) # proportion of sexes
addmargins(prop.table(table(train$Sex))) # plain proportion with control sum (1.000000)
```
```{r}
# check the total numbers of passed (=0) and survived (=1) by sex
table(train$Survived, train$Sex)
addmargins(prop.table(table(train$Survived, train$Sex),1))
# (prop.table -> ratio by row =1, by column =2)
```
```{r}
# assign "0" to all survived, except females getting "1"
test$Survived <- 0
test$Survived[test$Sex == 'female'] <- 1

# check the changes
head(test$Survived)
head(test$Sex)
table(test$Survived, test$Sex)
```
```{r}
# make the second submission with R, rate = 0.76555
submit <- data.frame(PassengerId = test$PassengerId, Survived = test$Survived)
write.csv(submit, file = "theypartiallygone.csv", row.names = FALSE)
```

Getting the age info
```{r}
summary(train$Age)
```

```{r}
# create a Child column for passengers aged under 18
train$Child <- 0
train$Child[train$Age < 18] <- 1
str(train$Child)
```
```{r}
# get the general digits for Child appearance in train data
aggregate(Survived ~ Child, data = train, FUN = sum)
```

